\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-06-09 doc/paper.tex
doc/paper.tex: fix typos

To study the properties of SUSY models programs are needed which
numerically calculate the pole masses and couplings of the SUSY
particles given a set of theory input parameters.  The output of these
so-called spectrum generators can be transferred to programs which
calculate further observables such as branching rations or the dark
matter relic density.

@@

Using the the variable \code{EWSBOutputParameters} the user can
specify the model parameters that are output of the electroweak
symmetry breaking consistency conditions.  When imposing the EWSB, \fs
will adjust these parameters until the EWSB conditions are fulfilled.
In the CMSSM example above these are the superpotential parameter
$\mu$ and its corresponding soft-breaking parameter $B\mu$.  In the
NMSSM the parameters $\kappa$, $|v_s|$ and $m_s^2$ are usually chosen
for this purpose.

@@

\item The function \code{FSMinimize[parameters, function]} can be
  given, where \code{parameters} is a list of model parameters and
  \code{function} is a function of these parameters.
  \code{FSMinimize[parameters, function]} will numerical vary the
  \code{parameters} until the \code{function} is minimized.  Example:
  %
  \begin{lstlisting}
FSMinimize[{vd,vu},
           (SM[MZ] - Pole[M[VZ]])^2 / STANDARDDEVIATION[MZ]^2 +
           (SM[MH] - Pole[M[hh[1]]])^2 / STANDARDDEVIATION[MH]^2]
  \end{lstlisting}
  %
  Here, the parameters \code{vu} and \code{vd} are varied until the
  function
  %
  \begin{align}
    \chi^2(v_d,v_u) =
    \frac{(\texttt{SM[MZ]}-m_Z^\pole)^2}{\sigma_{m_Z}^2} +
    \frac{(\texttt{SM[MH]}-m_{h_1}^\pole)^2}{\sigma_{m_h}^2}
  \end{align}
  %
  is minimal.  The constants \code{SM[MZ]}, \code{SM[MH]},
  $\sigma_{m_Z}$ and $\sigma_{m_h}$ are defined in
  \code{src/ew_input.hpp} to be
  %
  \begin{align}
    \texttt{SM[MZ]} &= 91.1876, &
    \texttt{SM[MH]} &= 125.9, \\
    \sigma_{m_Z} &= 0.0021, &
    \sigma_{m_h} &= 0.4 .
  \end{align}

@@

\fs allows to add leading two-loop contributions to the CP-even Higgs
tadpoles and self-energies.  For MSSM-like models (with two CP-even
Higgs bosons, one CP-odd Higgs boson, one neutral Goldstone boson)
these corrections can be enabled by setting \code{UseHiggs2LoopMSSM =
  True} in the model file and by defining the effective $\mu$-term
\code{EffectiveMu = \\[Mu]}.  This will add the zero-momentum
corrections of the order $O(y_t^4 + y_b^2 y_t^2 + y_b^4)$, $O(y_t^2
g_3^2)$, $O(y_b^2 g_3^2)$, $O(y_\tau^4)$ from
\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}.
For NMSSM-like models (with three CP-even Higgs bosons, two CP-odd
Higgs bosons, one neutral Goldstone boson) the two-loop contributions
are enabled by setting \code{UseHiggs2LoopNMSSM = True} and by
defining the effective $\mu$-term like \code{EffectiveMu = \\[Lambda]
  vS / Sqrt[2]}, for example.  This will add the the zero-momentum
corrections of the order $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$ from
\cite{Degrassi:2009yq}, plus MSSM-like contributions of the order
$O(y_\tau^4)$, $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$
\cite{Brignole:2001jy,Dedes:2003km}.

@@

\item The \emph{low-scale constraint} is intended to match the SUSY
  model to the Standard Model at the scale $M_Z$.  It does so by
  calculating the gauge couplings of the SUSY model from the known
  Standard Model quantities
  $\alpha_{\text{e.m.},\text{SM}}^{(5),\text{\MSbar}}(M_Z)$
  $\alpha_{\text{s},\text{SM}}^{(5),\text{\MSbar}}(M_Z)$, $M_Z$ and
  $M_W$.  This calculation is explained in
  \secref{sec:calculation-of-gauge-couplings}.  Optionally, the Yukawa
  couplings of the SUSY model can be calculated automatically from the
  Standard Model fermion masses.  See
  \secref{sec:calculation-of-yukawa-couplings} for more details.  In
  addition to the gauge and Yukawa couplings, the model parameter
  constraints given in \code{LowScaleInput} are imposed here.
\end{itemize}

@@

\subsection{Pole masses}
\label{sec:PoleMasses}
After the solver routine has finished and convergence has been
achieved, all \DRbar\ parameters consistent with the EWSB conditions,
low energy data and all user-supplied boundary conditions and are
known at any scale between \code{LowScale} and \code{HighScale}.

@@

For the run-time comparison we're generating $2\cdot 10^{4}$ random
CMSSM parameter points with $m_0\in [50,1000]\unit{GeV}$, $m_{1/2}\in
[50,1000]\unit{GeV}$, $\tan\beta\in [1,100]$, $\sign\mu\in \{-1,+1\}$
and $A_0\in [-1000,1000]\unit{GeV}$.  For each point an SLHA input
file is created by appending the values of $m_0$, $m_{1/2}$,
$\tan\beta$, $\sign\mu$, $A_0$ in form of a \code{MINPAR} block to the
SLHA template file given in
\secref{sec:speed-test-slha-template-file}.  The resulting SLHA input
file is passed to each spectrum generator and the (wall-clock) time is
measured until the program has finished.  The average run-times for
three different CPU types can be found in
\tabref{tab:run-time-comparison}.  The first column shows the run-time
on a Intel Core2 Duo (P8600, $2.40\unit{GHz}$) where only one core was
enabled.  The second column shows the run-time on the same processor
where both cores were enabled.  In the third column a Intel Xeon
(L5640, $2.27\unit{GHz}$) was used, which has $6$ CPU cores.
%
\begin{table}[tbh]
  \centering
  \begin{tabular}{llll}
    \toprule
                            & Intel Core2 Duo    & Intel Core2 Duo   & Intel Xeon\\
                            & (P8600, 1 core)    & (P8600, 2 cores)  & (L5640, 6 cores)\\
    \midrule
    FlexibleSUSY-NoFV 1.0.0 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
    SPheno 3.2.4            & $0.119\unit{s}$    & $0.114\unit{s}$   & $0.101\unit{s}$\\
    Softsusy 3.4.0          & $0.175\unit{s}$    & $0.171\unit{s}$   & $0.147\unit{s}$\\
    \midrule
    FlexibleSUSY-FV 1.0.0   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
    SPhenoMSSM 4.1.0        & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
    \bottomrule
  \end{tabular}
  \caption{Average run-time of CMSSM spectrum generators for
    for random parameter points.  The first three rows show
    spectrum generators which disallow flavour violation.  Rows
    $4$--$5$ contain flavour violating spectrum generators, based
    on SARAH's MSSM model file.}
  \label{tab:run-time-comparison}
\end{table}

>>>>

To study the properties of SUSY models programs are needed which
numerically calculate the pole masses and couplings of the SUSY
particles given a set of theory input parameters.  The output of these
so-called spectrum generators can be transferred to programs which
calculate further observables such as branching ratios or the dark
matter relic density.

@@

Using the variable \code{EWSBOutputParameters} the user can
specify the model parameters that are output of the electroweak
symmetry breaking consistency conditions.  When imposing the EWSB, \fs
will adjust these parameters until the EWSB conditions are fulfilled.
In the CMSSM example above these are the superpotential parameter
$\mu$ and its corresponding soft-breaking parameter $B\mu$.  In the
NMSSM the parameters $\kappa$, $|v_s|$ and $m_s^2$ are usually chosen
for this purpose.

@@

\item The function \code{FSMinimize[parameters, function]} can be
  given, where \code{parameters} is a list of model parameters and
  \code{function} is a function of these parameters.
  \code{FSMinimize[parameters, function]} will numerically vary the
  \code{parameters} until the \code{function} is minimized.  Example:
  %
  \begin{lstlisting}
FSMinimize[{vd,vu},
           (SM[MZ] - Pole[M[VZ]])^2 / STANDARDDEVIATION[MZ]^2 +
           (SM[MH] - Pole[M[hh[1]]])^2 / STANDARDDEVIATION[MH]^2]
  \end{lstlisting}
  %
  Here, the parameters \code{vu} and \code{vd} are varied until the
  function
  %
  \begin{align}
    \chi^2(v_d,v_u) =
    \frac{(\texttt{SM[MZ]}-m_Z^\pole)^2}{\sigma_{m_Z}^2} +
    \frac{(\texttt{SM[MH]}-m_{h_1}^\pole)^2}{\sigma_{m_h}^2}
  \end{align}
  %
  is minimal.  The constants \code{SM[MZ]}, \code{SM[MH]},
  $\sigma_{m_Z}$ and $\sigma_{m_h}$ are defined in
  \code{src/ew_input.hpp} to be
  %
  \begin{align}
    \texttt{SM[MZ]} &= 91.1876, &
    \texttt{SM[MH]} &= 125.9, \\
    \sigma_{m_Z} &= 0.0021, &
    \sigma_{m_h} &= 0.4 .
  \end{align}

@@

\fs allows to add leading two-loop contributions to the CP-even Higgs
tadpoles and self-energies.  For MSSM-like models (with two CP-even
Higgs bosons, one CP-odd Higgs boson, one neutral Goldstone boson)
these corrections can be enabled by setting \code{UseHiggs2LoopMSSM =
  True} in the model file and by defining the effective $\mu$-term
\code{EffectiveMu = \\[Mu]}.  This will add the zero-momentum
corrections of the order $O(y_t^4 + y_b^2 y_t^2 + y_b^4)$, $O(y_t^2
g_3^2)$, $O(y_b^2 g_3^2)$, $O(y_\tau^4)$ from
\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}.
For NMSSM-like models (with three CP-even Higgs bosons, two CP-odd
Higgs bosons, one neutral Goldstone boson) the two-loop contributions
are enabled by setting \code{UseHiggs2LoopNMSSM = True} and by
defining the effective $\mu$-term like \code{EffectiveMu = \\[Lambda]
  vS / Sqrt[2]}, for example.  This will add the zero-momentum
corrections of the order $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$ from
\cite{Degrassi:2009yq}, plus MSSM-like contributions of the order
$O(y_\tau^4)$, $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$
\cite{Brignole:2001jy,Dedes:2003km}.

@@

\item The \emph{low-scale constraint} is intended to match the SUSY
  model to the Standard Model at the scale $M_Z$.  It does so by
  calculating the gauge couplings of the SUSY model from the known
  Standard Model quantities
  $\alpha_{\text{e.m.},\text{SM}}^{(5),\text{\MSbar}}(M_Z)$,
  $\alpha_{\text{s},\text{SM}}^{(5),\text{\MSbar}}(M_Z)$, $M_Z$ and
  $M_W$.  This calculation is explained in
  \secref{sec:calculation-of-gauge-couplings}.  Optionally, the Yukawa
  couplings of the SUSY model can be calculated automatically from the
  Standard Model fermion masses.  See
  \secref{sec:calculation-of-yukawa-couplings} for more details.  In
  addition to the gauge and Yukawa couplings, the model parameter
  constraints given in \code{LowScaleInput} are imposed here.
\end{itemize}

@@

\subsection{Pole masses}
\label{sec:PoleMasses}
After the solver routine has finished and convergence has been
achieved, all \DRbar\ parameters consistent with the EWSB conditions,
low energy data and all user-supplied boundary conditions are
known at any scale between \code{LowScale} and \code{HighScale}.

@@

For the run-time comparison we're generating $2\cdot 10^{4}$ random
CMSSM parameter points with $m_0\in [50,1000]\unit{GeV}$, $m_{1/2}\in
[50,1000]\unit{GeV}$, $\tan\beta\in [1,100]$, $\sign\mu\in \{-1,+1\}$
and $A_0\in [-1000,1000]\unit{GeV}$.  For each point an SLHA input
file is created by appending the values of $m_0$, $m_{1/2}$,
$\tan\beta$, $\sign\mu$, $A_0$ in form of a \code{MINPAR} block to the
SLHA template file given in
\secref{sec:speed-test-slha-template-file}.  The resulting SLHA input
file is passed to each spectrum generator and the (wall-clock) time is
measured until the program has finished.  The average run-times for
three different CPU types can be found in
\tabref{tab:run-time-comparison}.  The first column shows the run-time
on a Intel Core2 Duo (P8600, $2.40\unit{GHz}$) where only one core was
enabled.  The second column shows the run-time on the same processor
where both cores were enabled.  In the third column a Intel Xeon
(L5640, $2.27\unit{GHz}$) was used, which has $6$ CPU cores.
%
\begin{table}[tbh]
  \centering
  \begin{tabular}{llll}
    \toprule
                            & Intel Core2 Duo    & Intel Core2 Duo   & Intel Xeon\\
                            & (P8600, 1 core)    & (P8600, 2 cores)  & (L5640, 6 cores)\\
    \midrule
    FlexibleSUSY-NoFV 1.0.0 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
    SPheno 3.2.4            & $0.119\unit{s}$    & $0.114\unit{s}$   & $0.101\unit{s}$\\
    Softsusy 3.4.0          & $0.175\unit{s}$    & $0.171\unit{s}$   & $0.147\unit{s}$\\
    \midrule
    FlexibleSUSY-FV 1.0.0   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
    SPhenoMSSM 4.1.0        & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
    \bottomrule
  \end{tabular}
  \caption{Average run-time of CMSSM spectrum generators
    for random parameter points.  The first three rows show
    spectrum generators which disallow flavour violation.  Rows
    $4$--$5$ contain flavour violating spectrum generators, based
    on SARAH's MSSM model file.}
  \label{tab:run-time-comparison}
\end{table}
